The importance of measurement uncertainty in ecological management

### 2018-01-04 manuscripts/manuscript.Rmd
WIP

abstract: |
  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to the intuition of the precautionary principle. Through the use of cutting-edge point-based POMDP methods, we demonstrate that the standard management theory which ignores measurement uncertainty may lead to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, POMDP solutions which account for measurement error are overall more conservative, avoiding over-exploitation while generating higher economic value.

@@

```{r suggested_libs}
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
```

@@

The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

@@

We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.  To facilitate tractability and intepretation, we will consider a simple and well-studied model of population growth (stock recruitment) in marine fisheries, the Graham-Schaefer model:

Though this and very similar models are widely used in large scale analyses across diverse stocks (e.g. [@Costello2016; @Britten2017]), and common in other conservation contexts where richer models are unavailable, neither the existing approaches nor the one we introduce here are limited to such one-dimensional stochastic models with known parameters. In practice, this model may include age or stage structure and will reflect uncertainty not only in the stochastic recruitment but in the parameters themselves, usually as the result of a heirarchical model estimation process.   All that is required of the model is that it defines the probability of transitioning to a new state given the current state (here, simply the stock size) and the action taken (here, just the harvest).  Uncertainty either in the parameters, in the model, or inherent in the stochastic process can all be reflected in such probabilities.

Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

$$y_t = Z_m x_t$$
for some measurement error $Z_m$.  Given the estimated recruitment model along with this estimate of the stock size, the manager must set a harvest quota $h_t$ for that year.  The methods we consider here all differ primarily in their handling of these uncertainties. If this system were completely deterministic so that there was no uncertainty, all would produce identical long term policies and thus achieve identical outcomes.

In the case of no uncertainty, the optimal harvest strategy is to reduce the population to the size at which it obtains the maximum growth rate, known as Biomass at Maximum Sustainable Yield, $B_{MSY}$ (this is adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990). For the Graham-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$, with a corresponding annual harvest of $H_{MSY} = r K / 4$.

This equilibrium analysis says nothig about the appropriate harvest at stock sizes above or below $B_{MSY}$, which highlights some essential differences between the optimal control policy and more typical fisheries management.  The optimal policy is impatient: it seeks to achieve the optimal yield as quickly as possible.  Below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, a harvest greater than $H_{MSY}$ can be optimal for that year.  This strategy is known as "constant escapement," since a constant sock size $B_{MSY}$ escapes harvest each year.  In practice, it is more common to pursue a policy of constant fishing effort (fishing mortality), $F_{MSY} = H_{MSY} / B_{MSY}$.  Continuing to harvest at $F_{MSY}$ when a stock is above or below $B_{MSY}$ is economically inefficient and causes depleted stocks to rebuild more slowly, but, in the absense of uncertainty, ensures a return in the long run to stock sizes of $B_{MSY}$ and harvests of $H_{MSY}$.  These differences will be important in understanding how these approaches perform once uncertainty is introduced.

@@

As discussed above, absent the short-cut provided by Reed's theorems, solving for the optimal management policy in a stochastically fluctuating population involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP). Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).  The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process.  The POMDP problem for fisheries question considered here can be summarized as follows:

>>>>

abstract: |
  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle. Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.

@@

```{r suggested_libs, message=FALSE}
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
```

@@

Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.

We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

@@

We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.  To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1974; @Gordon1974] of logistic population growth (stock recruitment),

$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$

where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.  We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.

This and similar models are widely used in large scale analyses across diverse stocks (e.g. [@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990]. As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.

However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.  Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g. @].  As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.

## The decision problem

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\delta$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

$$y_t = \xi_t x_t$$

Given an estimated recruitment model along with this measurement of the stock size, the manager must set a harvest quota $h_t$ for that year in order to maximize their expected long term (net) present value:

$$\max_{ \lbrace h_t \rbrace} \sum_{t=0}^{\infty} \delta x_t $$

A few subtlies arise in this seemlingly simple problem statement that we must address.  First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.

A second subtly is in the nature of the decision problem itself: the optimization problem must always consider the set ${h_t}$ all future actions simultaneously, not merely the next action.  This is a game of chess, where it can be optimal to make costly short term actions (such as closing the fishery) if they pay off in the long run. This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a single harvest quota $H_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.  Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.

## Current theory and current practice

@Gordon1954 and @Schaefer1954 independently showed in the same year that the maximum sustainable yield for the model bearing their names is achieved by reducing the stock to the  population to the size at which it obtains the maximum growth rate which is known as Biomass at Maximum Sustainable Yield, $B_{MSY}$. For the Gordon-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$.  They observed that fishing at a *constant yield* (that is, an individual fish mortality, $F$, such that harvest $H = F \cdot B$) of $F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}$)  will eventually lead to a population that converges to the biomass $B_{MSY}$ and produces the maximum sustainable harvest, $H_{MSY} = r K / 4$ in this model.
From here, theory and practice diverge.

This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis, and does not solve the time-dependent optimization problem above. In particular, maximum sustainable yield will not be optimal whenever the stock is away from $B_{MSY}$.  @Clark1973 shows that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*. The time-depednent optimal stategy is impatient: it seeks to achieve the optimal yield as quickly as possible.  Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, mortalities greater than $F_{MSY}$ can be optimal to bring the stock back to $B_{MSY}$ more quickly.  (This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.

@@

As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP). Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).  The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process.  The POMDP problem for fisheries question considered here can be summarized as follows:
